---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# densmix

<!-- badges: start -->
<!-- badges: end -->

The goal of densmix is to fit density mixture models to data.

Two mixture models are supported:

- One dimensional exponential - normal density.
- Mixture of multinomials.

## Installation

You can install the development version of densmix from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("agehsbargswork-blip/data-science")
```

## Example: exp-normal density

Here we generate exponential-normal data and check if algorithm reconstructs it: 

```{r example1}
library(densmix)

params <- list(
  weight=0.5,
  lambda=0.6,
  mu=10,
  sigma=2
)

data <- gen_exp_normal(size=1000,params)
emfit <- fit_exp_normal(data)
plot(emfit)

```

## Example: multinomial density

Here we generate multinomial data with three densities and check if algorithm reconstructs it. We compare all pairs of densities as order of reconstruction is arbitrary.

```{r example2}
library(densmix)
set.seed(123)
gendata <- gen_em_mn(size = 1000,
                     params = list("n_components" = 3,
                                   "n_buckets" = 12,
                                   "dirichlet" = rep(1,12),
                                   "component_weights" = c(0.1,0.2,0.7),
                                   "n_actions_per_bucket" = list(
                                     "max_actions" = 100,
                                     "prob_action" = 0.5
                                   )
                     )
      )


fitdata <- fit_multinomial(gendata$data, control = list(verbose = TRUE))

par(mfrow=c(3,3))
for(i in 1:3){
  for(j in 1:3){
    plot(gendata$mixture_profiles[i,]
         ,fitdata$mixture_profiles[j,]
         ,pch=19
         ,xlab=paste("Generated profile ",i,sep="")
         ,ylab=paste("Fitted profile ",j,sep="")
         )
  }
}

plot(fitdata)
```

